
# Chapter 1

## Questions

1. What are the properties of a vector other than its contents?
  - Three common properties
    - type `typeof()`   
    - length `length()`
    - attributes `attributes()`

2. What are the four common types of atomic vectors? What are the two rare types?
  - logical, double, character, integer (same as double with no decimal point)
   - complex, raw

3. What are attributes and how to you get and set them?

4. How is a list different from an atomic vector? How is a matrix different from a Data Frame?

5. Can you have a list that is a matrix? Can a Data Frame have a column that is a matrix?



### Q1 - what are the properties of a vector other than its contents?

The basic Data structure in R is a vector - they come in **Atomic Vectors** and **Lists**.


```{r}
typeof() ## what it is

length() ## How many elements it contains

attributes() ## additional arbitrary metadata
```

#### Atomic Vectors

[Vector Explaination](https://renenyffenegger.ch/notes/development/languages/R/data-structures/vector/)

##### Four kinds
- logical (true/ false)
- integer (numbers without decimals)
- double (numbers with decimals)
- character (strings/ words)
```{r}
dbl_var <- c(1,1.5,2.5) 

int_var <- c(1L, 2.5L, 5L) ## with the L suffic there is ain integer rather than a double

log_var <- c(TRUE, FALSE, TRUE, T, F)

chr_var <- c("these are some words", "these are some other words", "strings are 'words'.")

```

- Notes
  - Atomic vectors are always flat, even if you nest `c()`'s e.g `c(1,c(1,3.4))` is the same as `c(1,2,3.4)`
  - Missing values are replaced with `NA` which is a logical vector of '1'.

##### Types and Tests

```{r}
typeof(dbl_var)
```

```{r}
is.atomic(log_var)
```
```{r}
is.double(dbl_var)
```
```{r}
is_numeric(dbl_var) ## from purrr doing the same as is.numeric but is deprecated
is.numeric(dbl_var)
```
##### Coercion

All elements of a vector must be of the same type. If they aren't then they will be coerced into the most flexible type. 
```{r}
str(c(1,"a")) ## integers and strings combine to form 'character' vectors
```
 When a logical vector is coerced to integer or double, `TRUE = 1` and `FALSE = 0`. Often used in conjunction with `sum()` and `mean()`. 
```{r}
x <- c(TRUE, FALSE, TRUE, FALSE, FALSE, TRUE)
typeof(x)
sum(x) ## total number of 'TRUE's
as.numeric(x)
mean(x) ## Proportion that are 'TRUE'
```
 If there is any chance of confusion specifically coerce a vector to another type with
 - `as.logical`
 - `as.integer`
 - `as.character`
 - `as.double`
 
 --- 
#### Lists 

