


# Using dplyr

```{r}
library(tidyverse)
```


The [dlpyr website]https://dplyr.tidyverse.org) has much more detail than is in this workbook. This is the start of manipulation skills. I will endeavour to incorporate some more advanced R comments from [Advanced R](https://learning.oreilly.com/library/view/advanced-r/9781466586963/K20319_C003.xhtml) as we progress through the basics.

For the purposes of this work we will use a set of HR Data. The data hasn't not been cleaned at all so we wil go through the full process during this work cycle. 

The data can be viewed below

```{r}
head(X2010_csv)

view(X2010_csv)

## Tidy the Columns Names by removing the spaces and standardising the output

df <- X2010_csv %>%
  rename_with( ~ tolower(gsub(" ", "_", .x)))

df
```

An initial EDA reveals the following

```{r}
## Tidy the Data
df1 <- df %>%
  filter(!termination_reason == "Separation") %>% ## Remove the duplicate reasons for leaving
  select(!c(name,servop,subunit,unit, service_nbr, previous_subunit, commanding_off._posn))%>%  ## De-identify the Data and remove unecessary columns
  distinct(emplid, .keep_all = TRUE) ## Keep only the unique values within the DF 

df1

df2 <- df1[duplicated(df1$emplid),] ## This is a list of all the values in the DF that were not unique - 66 in total

df2


```

While a quick summary of the data doesn't reveal much the date columns are interesting. We can see that all the termination dates are in the year of 2010 while the Action dates range from 2005 to 2015. There is also some interesting differences in the service dates and the hire dates. 

```{r}
  
rand <- runif(6514) %>%
  as_tibble()%>%
  view()

rand

deidentify <- function(x) {
  rand <- round(runif(x)*x,2)
} ## Funciton to generate random numbers and multiply them by the original number - used to deidentify the emplyee id


df1 <- df1%>%
  mutate(
    emplid = deidentify(emplid)
  )

df1

```


```{r}
## Transform the ranks into Factors for further analysis

distinct(df1, rank)

rank_levels <- c(
  "PTE TRN", "PTE", "PTE(P)", "LCPL", "CPL", "SGT", "SSGT", "WO2", "WO1", "LT", "CAPT", "MAJ", "LTCOL", "COL", "BRIG", "PTE REC", "OCDT", "MAJGEN", "2LT")

distinct(df1, family)

corp_family <- c(
  "ASUT", "RAAC", "RAE", "RA SIGS", "RAINF", "RACT", "RAAMC", "RAAOC", "RAA", "AABC", "AUST INT", "AACC", "RAEME", "RACMP", "AAPRS", "AAAVN", "AAPSYCH", "RAANC", "RAADC", "AALC", "RAA CH D", "RAAEC", "SENOFF", "RA INF", "AA PSYCH", "Uknown", "NON CORPS", "LOCOBS")
)

distinct(df1, service)

service_types <- c("CFTS", "HRR CFTS", "AFS", "ARA", "REG-GAP", "AAR", "RFSL", "RES-HRR", "ASR", "RES-LOE","RES-RRF")

distinct(df1, termination_reason)

termination_type <- c("Military - Completed CFTS", "Military-Medical Unfit Service", "Military - Resignation", "Military - Contract Completed", "Military - Training Failure", "Mil-Within 90days of enlistmnt", "Mil-Retentn Not In Srvc Intrst", "Military-Below Fitness Standrd", "Military-Irregular enlistment", "Mil-False statemnt on enlistmt", "Mil-Unsuit For Svc DPR 87-1-e", "Military - In Absence", "Military - Death", "Military-Compulsory Retire Age", "Military-Mgt Initiate Retiremt", "Military - Disciplinary", "Military - Civil Offence", "Military -Retired After Age 55")

distinct(df1, previous_unit)

factor(termination_type)

## Transform categorical values levels into factors

df1 <- df1 %>% mutate(
  rank = factor(rank, levels = rank_levels),
  corps = factor(family, levels = corp_family),
  service = factor(service, levels = service_types),
  termination_reason = factor(termination_reason, levels = termination_type),
  previous_unit = as.factor(previous_unit)
)

str(df1)
head(df1)
glimpse(df1)
levels(df1$rank)

```

```{r}

df1 %>% ## pipe the data into a visualisation
  mutate(rank = rank %>% fct_infreq() %>% fct_rev())%>%   ## using pipes inside the mutate to order levels and reverse the order
  ggplot(aes(rank, fill = rank)) + 
  geom_bar() +
  coord_flip() +
  labs(title = "Total Separations by Rank", x = "Rank", y = "Total") +
  theme_minimal() 

```

```{r}
## sorting percentages of total terminations by rank

df1 %>% 
  group_by(rank) %>%
  summarise(percentage = n(), .groups = "drop") %>% 
  mutate(pct = round(percentage/sum(percentage)*100,2)) %>% 
  ggplot(aes(reorder(rank,pct, order = TRUE), pct, color = rank, )) +
    geom_point() +
    geom_text(aes(label = pct), nudge_y = 2) +
    coord_flip() +
    ggtitle("Total Separations by Rank - 2010") +
    theme_minimal() +
    labs(title = "Percentage of Total Separations by Rank", x = "Rank", y = "Percentage") +
    guides(color = FALSE) 
    
  


```


```{r}

## Plots trade and rank




```

